---
title: "Time-Series: Forecasting"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---


```{r}

library(tidyverse)
library(flexdashboard)
library(shiny)
library(shinyWidgets)
library(plotly)
library(tsibble)
library(tsibbledata)
library(feasts)
library(fable)
library(fabletools)
library(urca)
library(fable.prophet)

```



```{r}


rm(list=ls())

df<-fpp3::us_employment

df <- df |>
  filter(year(Month) >= 1990, Title == "Retail Trade") |>
  select(-Series_ID)|>
  mutate(diff = difference(Employed),
         log = log(Employed),
         diff_log = difference(log))



dcmp <- df |>
  model(stl = STL(Employed))


comps<-components(dcmp)


choices<-c("none","difference","log","diff_log")

```




<b>Decomposition</b> 
=====================================  


Sidebar {.sidebar}
-------------------------------------


```{r}

pickerInput(
  inputId = "transformation", 
  label = HTML('<b style="color:#234370;
               font-size: 20px;
               ">Transformation</b>'), 
  choices = choices,
  selected = "none"
)



```


**Transformation Filter**: Allows user to see different transformations to the series.

The model used for decomposition is the **STL** (Seasonal and Trend decomposition using Loess)


<style>

.chart-title {
    border-bottom: 1px solid #dfdfdf;
    font-size: 14px;
    font-weight: 300;
    padding: 7px 10px 4px;
    color: white;
    font-family: inherit;
    background-color: #234370;
}

</style>



<style>

.navbar-inverse {
    background-color: #2b2d38;
    border-color: #2b2d38;
}

</style>


<style>

.navbar, [data-toggle=tab], .navbar-brand  {
background-color:#2b2d38!important;
boarder-color: #fffff!important;}

</style>



<style>

.section.sidebar {
    top: 51px;
    background-color: #ffffff;
}

</style>












Outputs {data-width=650} 
-------------------------------------

### <b>US Retail Employment</b>

```{r}

renderPlotly({
 
  if(input$transformation=="none"){ 
  p<-autoplot(df, Employed) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  } else if(input$transformation=="difference"){
    
    p<-autoplot(df, diff) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  }else if(input$transformation=="log"){
    
    p<-autoplot(df, log) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  }else if(input$transformation=="diff_log"){
    
    p<-autoplot(df, diff_log) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  }
  
  ggplotly(p)

})

```

Column {data-width=650} 
-------------------------------------

### <b>Trend</b>


```{r}

renderPlotly({
  
  
  p<-autoplot(comps, trend) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  
  
  ggplotly(p)
  
})


```


### <b>Season Year</b>


```{r}

renderPlotly({
  
  
  p<-autoplot(comps, season_year) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  
  
  ggplotly(p)
  
})


```


### <b>Remainder</b>


```{r}

renderPlotly({
  
  
  p<-autoplot(comps, remainder) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  
  
  ggplotly(p)
  
})


```



<b>Prewhitening</b> {data-orientation=rows}
=====================================  

Sidebar {.sidebar}
-------------------------------------

```{r}

pickerInput(
  inputId = "transformation2", 
  label = HTML('<b style="color:#234370;
               font-size: 20px;
               ">Transformation</b>'), 
  choices = choices,
  selected = "none"
)



```


**Transformation Filter: Allows user to see different transformations to the series.**


<style>

.chart-title {
    border-bottom: 1px solid #dfdfdf;
    font-size: 14px;
    font-weight: 300;
    padding: 7px 10px 4px;
    color: white;
    font-family: inherit;
    background-color: #234370;
}

</style>



<style>

.navbar-inverse {
    background-color: #2b2d38;
    border-color: #2b2d38;
}

</style>


<style>

.navbar, [data-toggle=tab], .navbar-brand  {
background-color:#2b2d38!important;
boarder-color: #fffff!important;}

</style>



<style>

.section.sidebar {
    top: 51px;
    background-color: #ffffff;
}

</style>












Column 
-------------------------------------

### <b>Stationary Test<br>P-Value<br></b>
```{r}
cols<-c("Employed","diff","log","diff_log")


x<-map(df[,cols],unitroot_kpss)%>%as_tibble()




renderValueBox({
 
  if(input$transformation2=="none"){ 
  p<-x[2,"Employed"]
  
  } else if(input$transformation2=="difference"){
    
    p<-x[2,"diff"]
  } else if(input$transformation2=="log"){
    
    p<-x[2,"log"]
  }else if(input$transformation2=="diff_log"){
    
    p<-x[2,"diff_log"]
  }
  

    valueBox(p, 
           icon = ifelse(test = p < .05,
                        yes = "fa fa-thumbs-down",
                        no = "fa fa-thumbs-up"),
           color = ifelse(test = p < .05, 
                          yes = "danger", 
                          no = "primary"))

})




```



### <b>Average</b>
```{r}
cols<-c("Employed","diff","log","diff_log")


x<-map(df[,cols],unitroot_kpss)%>%as_tibble()
x1<-round(map(df[,cols],mean,na.rm=T)%>%as_tibble(),2)



renderValueBox({
 
  if(input$transformation2=="none"){ 
  p<-x[2,"Employed"]
  z<-x1[1,"Employed"]
  
  } else if(input$transformation2=="difference"){
    
    p<-x[2,"diff"]
    z<-x1[1,"diff"]
  } else if(input$transformation2=="log"){
    
    p<-x[2,"log"]
    z<-x1[1,"log"]
  }else if(input$transformation2=="diff_log"){
    
    p<-x[2,"diff_log"]
    z<-x1[1,"diff_log"]
  }
  

    valueBox(z, 
           icon = ifelse(test = p < .05,
                        yes = "fa fa-thumbs-down",
                        no = "fa fa-thumbs-up"),
           color = ifelse(test = p < .05, 
                          yes = "danger", 
                          no = "primary"))

})




```



### <b>Standard Deviation</b>
```{r}
cols<-c("Employed","diff","log","diff_log")


x<-map(df[,cols],unitroot_kpss)%>%as_tibble()
x2<-round(map(df[,cols],sd,na.rm=T)%>%as_tibble(),2)



renderValueBox({
 
  if(input$transformation2=="none"){ 
  p<-x[2,"Employed"]
  z<-x2[1,"Employed"]
  
  } else if(input$transformation2=="difference"){
    
    p<-x[2,"diff"]
    z<-x2[1,"diff"]
  } else if(input$transformation2=="log"){
    
    p<-x[2,"log"]
    z<-x2[1,"log"]
  }else if(input$transformation2=="diff_log"){
    
    p<-x[2,"diff_log"]
    z<-x2[1,"diff_log"]
  }
  

    valueBox(z, 
           icon = ifelse(test = p < .05,
                        yes = "fa fa-thumbs-down",
                        no = "fa fa-thumbs-up"),
           color = ifelse(test = p < .05, 
                          yes = "danger", 
                          no = "primary"))

})




```





Row 
-------------------------------------

### <b>US Retail Employment</b>

```{r}

renderPlotly({
 
  if(input$transformation2=="none"){ 
  p<-autoplot(df, Employed) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  } else if(input$transformation2=="difference"){
    
    p<-autoplot(df, diff) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  }else if(input$transformation2=="log"){
    
    p<-autoplot(df, log) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  }else if(input$transformation2=="diff_log"){
    
    p<-autoplot(df, diff_log) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  }
  
  ggplotly(p)

})

```



<b>Identification</b> {data-orientation=rows}
=====================================  

Sidebar {.sidebar}
-------------------------------------


```{r}

pickerInput(
  inputId = "transformation3", 
  label = HTML('<b style="color:#234370;
               font-size: 20px;
               ">Transformation</b>'), 
  choices = choices,
  selected = "none"
)




```


**Identification Process**: See how many AR's and MA's are needed for implementing ARIMA Forecast Model

Row 
-------------------------------------

### <b>US Retail Employment</b>

```{r}

renderPlotly({
 
  if(input$transformation3=="none"){ 
  p<-autoplot(df, Employed) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  } else if(input$transformation3=="difference"){
    
    p<-autoplot(df, diff) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  }else if(input$transformation3=="log"){
    
    p<-autoplot(df, log) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  }else if(input$transformation3=="diff_log"){
    
    p<-autoplot(df, diff_log) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
  }
  
  ggplotly(p)

})

```


Column 
-------------------------------------

### <b>ACF</b>

```{r}

renderPlotly({
 
  if(input$transformation3=="none"){ 
    
    p<-df |> ACF(Employed) |> autoplot() +
      labs(y = "Persons (thousands)",
         title = "Total employment in US retail")
    
    
  } else if(input$transformation3=="difference"){
    
     p<-df |> ACF(diff) |> autoplot() +
      labs(y = "Persons (thousands)",
         title = "Total employment in US retail")
     
     
  }else if(input$transformation3=="log"){
    
     p<-df |> ACF(log) |> autoplot() +
      labs(y = "Persons (thousands)",
         title = "Total employment in US retail")
     
     
  }else if(input$transformation3=="diff_log"){
    
    p<-df |> ACF(diff_log) |> autoplot() +
      labs(y = "Persons (thousands)",
         title = "Total employment in US retail")
  }
  
  ggplotly(p)

})

```



### <b>PACF</b>

```{r}


renderPlotly({
 
  if(input$transformation3=="none"){ 
    
    p<-df |> PACF(Employed) |> autoplot() +
      labs(y = "Persons (thousands)",
         title = "Total employment in US retail")
    
    
  } else if(input$transformation3=="difference"){
    
     p<-df |> PACF(diff) |> autoplot() +
      labs(y = "Persons (thousands)",
         title = "Total employment in US retail")
     
     
  }else if(input$transformation3=="log"){
    
     p<-df |> PACF(log) |> autoplot() +
      labs(y = "Persons (thousands)",
         title = "Total employment in US retail")
     
     
  }else if(input$transformation3=="diff_log"){
    
    p<-df |> PACF(diff_log) |> autoplot() +
      labs(y = "Persons (thousands)",
         title = "Total employment in US retail")
  }
  
  ggplotly(p)

})

```



<b>Model</b> {data-orientation=rows}
===================================== 

Sidebar {.sidebar}
-------------------------------------

```{r}

numericInput("p",value = 0,min = 0,max = 12, step = 1,label = "ARIMA Input: P")

numericInput("d",value = 0,min = 0,max = 12, step = 1,label = "ARIMA Input: D")

numericInput("q",value = 0,min = 0,max = 12, step = 1,label = "ARIMA Input: Q")

```






Row 
-------------------------------------

### <b>ARIMA</b>

```{r}

renderDataTable({

  fit<-df|>
    model(arima013 = ARIMA(Employed ~ pdq(0,1,3)),
          arima_input = ARIMA(Employed  ~ pdq(input$p,input$d,input$q)),
          stepwise = ARIMA(Employed),
          search = ARIMA(Employed , stepwise=FALSE))

  glance(fit) |> arrange(AICc) |> select(.model:BIC)
  

})

```


### <b>Residual Diagnostics</b>

```{r}

renderPlot({
  
  resids<-df|>
    model(arima013 = ARIMA(Employed ~ pdq(0,1,3)),
          arima_input = ARIMA(Employed  ~ pdq(input$p,input$d,input$q)),
          stepwise = ARIMA(Employed),
          search = ARIMA(Employed , stepwise=FALSE))

  test<-glance(resids) |> arrange(AICc) |> select(.model:BIC)
  
  model<-test[1,1]


  text<-model$.model


  z<-resids[,text]|>
          gg_tsresiduals()
  
  
  z

})


```


Row 
-------------------------------------

### <b> Ljung-Box Test</b>

```{r}

renderDataTable({
  
  resids<-df|>
    model(arima013 = ARIMA(Employed ~ pdq(0,1,3)),
          arima_input = ARIMA(Employed  ~ pdq(input$p,input$d,input$q)),
          stepwise = ARIMA(Employed),
          search = ARIMA(Employed , stepwise=FALSE))

  test<-glance(resids) |> arrange(AICc) |> select(.model:BIC)
  
  model<-test[1,1]
  
  
  text<-model$.model
  
  
  augment(resids[,text]) |> features(.innov, ljung_box, lag=10)
  
})

```



<b>ARIMA Forecast</b> {data-orientation=rows}
===================================== 

Sidebar {.sidebar}
-------------------------------------

```{r}

sliderTextInput(
  inputId = "id", 
  label = HTML('<b style="color:#234370;
               font-size: 20px;
               ">Forecast Horizon:</b>'), 
  grid = TRUE, 
  force_edges = TRUE,
  choices = 1:50,
  selected = 5
  
)

```


Row
---------------------------------

### <b>US Retail Employment</b>

```{r}

renderPlot({
  
  fit<-df|>
      model(arima013 = ARIMA(Employed ~ pdq(0,1,3)),
            arima310 = ARIMA(Employed  ~ pdq(3,1,0)),
            stepwise = ARIMA(Employed),
            search = ARIMA(Employed , stepwise=FALSE))
  
  test<-glance(fit) |> arrange(AICc) |> select(.model:BIC)
  
  model<-test[1,1]
  
  
  text<-model$.model
  
  p<-fit |>
    forecast(h=input$id) |>
    filter(.model==text) |>
    autoplot(df)
  
  p
  
  
})


```



<b>Comparison Forecast</b> {data-orientation=rows}
===================================== 

Sidebar {.sidebar}
-------------------------------------

```{r}

sliderTextInput(
  inputId = "id2", 
  label = HTML('<b style="color:#234370;
               font-size: 20px;
               ">Forecast Horizon:</b>'), 
  grid = TRUE, 
  force_edges = TRUE,
  choices = 1:60,
  selected = 5
  
)

```


Row
---------------------------------

### <b>Comparison Forecast</b>

```{r}

renderPlot({
  
    train<- df|>
       filter(Month <= yearmonth("2015 Jan"))
    
    fit <- train |>
        model(
          arima = ARIMA(Employed),
          ets = ETS(Employed),
          prophet = prophet(Employed ~ season(period = 12, order = 10,
                                          type = "multiplicative"))
        )
  
    fc <- fit |> forecast(h = input$id2)|> autoplot(df)
    
    fc
    
    
})


```



Row 
-------------------------------------

### <b>Forecast Accuracy</b>

```{r}

renderDataTable({

    train<- df|>
           filter(Month <= yearmonth("2015 Jan"))
        
    fit <- train |>
      model(
        arima = ARIMA(Employed),
        ets = ETS(Employed),
        prophet = prophet(Employed ~ season(period = 12, order = 10,
                                        type = "multiplicative")))
      
    fc <- fit |> forecast(h = input$id2)
    
    fc |> accuracy(df)

})

```

